\documentclass[12pt, letterpaper]{memoir}
\usepackage{NotesStyle}

\begin{document}
	\begin{center}
		{\Large CHEM 3620}
		{\large-- Spring 2017
		
		Entropy and Irreversible Processes Quiz}
	\end{center}
		
	\subsection*{Problem 1 (5 points)}
Consider a spring which obeys Hook's law: $F=-kx$ where $x$ is the displacement away from equilibrium and $k=650\dfrac{N}{m}$. The acceleration due to gravity is $9.80665~\dfrac{m}{s^2}$.

$\circ$ Calculate the equilibrium displacement if a $10~kg$ weight is placed on the spring

\noindent 

\vspace{10em}
	
	\subsection*{Problem 2  (5 points)}	
Considering the same weight-on-a-spring in Problem 1:

$\circ$ Calculate the work done by the falling weight.

\vspace{10em}

	\subsection*{Problem 3 (5 points)}
How much work would be done if instead the spring was stretched reversibly to the same equilibrium displacement. Bonus -- Explain the discrepancy!

\vspace{10em}

	\subsection*{Problem 4 (5 points)}
Consider a system of $10.0~g$ of water super-cooled to $- 5^\circ C$ ($5^\circ C$ below the freezing point). The water suddenly freezes in an irreversible phase change. You will need to know the following values: $C_{V,water}=4.18\dfrac{J}{g^\circ C}$, $C_{V,ice}=2.03\dfrac{J}{g^\circ C}$, and $\Delta H_{fus, 0^\circ C} = 333.55 \dfrac{J}{g}$

$\circ$ Find $\Delta S_{system}$, $\Delta S_{surroundings}$, and $\Delta S_{universe}$ for this process.

(hint) You must find a fully reversible process with the same beginning and ending system state. To undergo this phase change reversibly, the system must be at the melting temperature.
\end{document}
	