\documentclass[11pt, letterpaper]{memoir}
\usepackage{HomeworkStyle}
\geometry{margin=1in}



\begin{document}

	\begin{center}
		{\large Quiz 17.2 -- The Arrhenius Equation}
	\end{center}
	{\large Name: \rule[-1mm]{4in}{.1pt} 


A reaction's rate is measured at different temperatures. Data are summarized in the table below:

\begin{tabular}{ccc}	
	Trial & Rate $\left(\nicefrac{M}{s}\right)$ & T $\left(^\circ C\right)$ \\ \midrule \midrule
	1 &$1.4\times10^{-4}$&$25$\\ \midrule
	2 &$3.7\times10^{-4}$&$75$ \\ \midrule
\end{tabular}

\noindent Use these data to find the activation energy for this reaction

\vspace{8em}
\noindent What would you expect the reaction rate to be at $-10^\circ C$?

\vspace{8em}
\noindent What is the Arrhenius pre-factor or frequency factor, $A$, for this reaction?


\vspace{8em}
\noindent The Arrhenius equation is based on a fairly simple model of reaction rates. What are some real factors which might make reaction rates deviate from the Arrhenius equation?

\newpage
\pagestyle{empty}
\addtocounter{page}{-1}	
\newgeometry{margin=1.25in}
\section*{\emph{Good Bones}}
\paragraph{By Maggie Smith}~
\begin{verse}
	Life is short, though I keep this from my children.\\
	Life is short, and I’ve shortened mine\\
	in a thousand delicious, ill-advised ways,\\
	a thousand deliciously ill-advised ways\\
	I’ll keep from my children. The world is at least\\
	fifty percent terrible, and that’s a conservative\\
	estimate, though I keep this from my children.\\
	For every bird there is a stone thrown at a bird.\\
	For every loved child, a child broken, bagged,\\
	sunk in a lake. Life is short and the world\\
	is at least half terrible, and for every kind\\
	stranger, there is one who would break you,\\
	though I keep this from my children. I am trying\\
	to sell them the world. Any decent realtor,\\
	walking you through a real shithole, chirps on\\
	about good bones: This place could be beautiful,\\
	right? You could make this place beautiful.
\end{verse}


\end{document}
